---
title: "10MinChallenge"
author: "Shane Regel"
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE, warnings = FALSE, messgae = FALSE)
```




## Load Data and Clean

Load Statcast data:
https://baseballsavant.mlb.com/leaderboard/custom?year=2019,2018,2017,2016,2015&type=batter&filter=&sort=4&sortDir=desc&min=100&selections=player_age,b_ab,b_total_pa,b_total_hits,b_single,b_double,b_triple,b_home_run,b_strikeout,b_walk,b_k_percent,b_bb_percent,batting_avg,slg_percent,on_base_percent,on_base_plus_slg,xba,xslg,woba,xwoba,xobp,xiso,exit_velocity_avg,launch_angle_avg,sweet_spot_percent,barrel_batted_rate,&chart=false&x=xba&y=xba&r=no&chartType=beeswarme

```{r}
library(tidyverse)
bat <- read_csv("stats.csv")
summary(bat)
```

Import MLB salary data from 2015-2019
https://www.usatoday.com/sports/mlb/salaries/2015/player/all/

```{r}
salaries <- read_csv("MLBsalaries.csv")
summary(salaries)
```

Split name column into first and last

```{r}
names <- data.frame(do.call('rbind', strsplit(as.character(salaries$NAME),' ',fixed=TRUE)))
names <- names[-c(3, 4)]
names(names) <- c("First", "Last")
salaries <- merge(salaries, names, by.x = 0, by.y = 0)
```

Drop unnecessary columns and pitchers from salary data

```{r}
salaries <- salaries[-c(1, 2, 3, 7, 8, 9)]
salaries <- salaries[!(salaries$POS=="SP" | salaries$POS=="RP"),]
```

Merge salary data with statcast data

```{r}
bat <- merge(bat, salaries, by.x=c("first_name", "last_name", "year"), by.y=c("First", "Last", "YEAR"))
bat <- bat[-c(32)]
```




## VISUALIZE THE DATA

Plot exit velocity and launch angle over time since statcasts release

```{r}
ggplot(bat, aes(x = year, y = exit_velocity_avg, group = year)) +            
  geom_boxplot() +
    labs(x = "Year", y = "Exit Velocity (mph)")
ggplot(bat, aes(x = year, y = launch_angle_avg, group = year)) +            
  geom_boxplot() +
    labs(x = "Year", y = "Launch Angle")
```

```{r}
ggplot(bat, aes(x = SALARY, y = exit_velocity_avg, scientific = FALSE)) +
    geom_point(aes(color = factor(player_age))) +
    labs(x = "Salary ($)", y = "Exit Velocity (mph)") +
    scale_x_continuous(labels = comma) +
    labs(colour = "Age")
ggplot(bat, aes(x = SALARY, y = launch_angle_avg)) +
    geom_point(aes(color = factor(player_age))) +
    labs(x = "Salary ($)", y = "Average Launch Angle") +
    scale_x_continuous(labels = comma) +
    labs(colour = "Age")
```

We see a lot of young players around the minimum salary line. This is due to the structure of rookie contracts in MLB. Lets strip those players out.

```{r}
nomin.lm.salary <- bat[!(bat$player_age<30 & bat$SALARY<555000),]
```

Now re-visualize

```{r}
ggplot(nomin.lm.salary, aes(x = SALARY, y = exit_velocity_avg, scientific = FALSE)) +
    geom_point(aes(color = factor(player_age))) +
    labs(x = "Salary ($)", y = "Exit Velocity (mph)") +
    scale_x_continuous(labels = comma) +
    labs(colour = "Age")
ggplot(nomin.lm.salary, aes(x = SALARY, y = launch_angle_avg)) +
    geom_point(aes(color = factor(player_age))) +
    labs(x = "Salary ($)", y = "Average Launch Angle") +
    scale_x_continuous(labels = comma) +
    labs(colour = "Age")
```
 



## LINEAR REGRESSION

Drop first name, last name, year, team, and position

```{r}
nomin.lm.salary <- nomin.lm.salary[-c(1, 2, 3, 32, 33)]
```

Run linear regression

```{r}
lm.nomin.salary <- lm(SALARY~., data=nomin.lm.salary)
summary(lm.nomin.salary)
confint(lm.nomin.salary)
```

Find the importance of each variable by training an lm model on the data

```{r}
library(caret)
lm.train.salary <- train(SALARY~., data=nomin.lm.salary, method="lm")
ggplot(varImp(lm.train.salary)) +
    labs(x = "Variable", y = "Importance")
```



## KNN
Turn salary into categorical variable
7 bins
$0 - $999,999
$1,000,000 - $4,999,999
$5,000,000 - $9,999,999
$10,000,000 - $14,999,999
$15,000,000 - $19,999,999
$20,000,000 - $24,999,999
$25,000,000 +

```{r}
library(FNN)
knn.salary <- nomin.lm.salary
knn.salary$SALARY <- cut(knn.salary$SALARY, breaks=c(0,1000000,5000000, 10000000, 15000000, 20000000, 25000000,100000000), right=FALSE, labels=c(1:7))
```

Plot distribution of the new salary bins

```{r}
ggplot() + 
  geom_histogram(aes(knn.salary$SALARY), bins=7, stat="count", color="darkblue", fill="lightblue") +
  labs(x = "Salary Bins ($ millions)", y = "Number of Players") +
  scale_x_discrete(labels=c("<1", "1-5", "5-10", "10-15", "15-20", "20-25", "25+"))
```
Create training and test data

```{r}
set.seed(111)
train.index <- sample(row.names(knn.salary), 0.6*dim(knn.salary)[1]) # 60% rows in training data 
valid.index <- setdiff(row.names(knn.salary), train.index)  # Rest of rows in test data
```

```{r}
train.df <- knn.salary[train.index, ]
valid.df <- knn.salary[valid.index, ]
```

KNN on the whole dataset with no minimum salaries, k=6

```{r}
nomin.knn.pred <- knn(train.df[, 1:28], valid.df[, 1:28], 
                cl = train.df[, 29], k = 6)

confusionMatrix(nomin.knn.pred, as.factor(valid.df[, 29]))
```

KNN on just exit velocity and launch angle for data with no minimum salaries, k=13

```{r}
nomin.knn.pred <- knn(train.df[, c(25,26)], valid.df[, c(25,26)], 
                cl = train.df[, 29], k = 13)

confusionMatrix(nomin.knn.pred, as.factor(valid.df[, 29]))
```

KNN on all variables except exit velocity and launch angle for data with no minimum salaries, k=10

```{r}
nomin.knn.pred <- knn(train.df[, c(1:24,27:28)], valid.df[, c(1:24,27:28)], 
                cl = train.df[, 29], k = 10)

confusionMatrix(nomin.knn.pred, as.factor(valid.df[, 29]))
```

Find most accurate K for all variables on no minimum salary dataset

```{r}
# initialize a data frame with two columns: k, and accuracy.
accuracy.df <- data.frame(k = seq(1, 14, 1), accuracy = rep(0, 14))

# compute knn for different k on validation.
for(i in 1:15) {
  nomin.knn.pred <- knn(train.df[, 1:28], valid.df[, 1:28], 
                  cl = train.df[, 29], k = i)
  accuracy.df[i, 2] <- confusionMatrix(nomin.knn.pred, as.factor(valid.df[, 29]))$overall[1] 
}

accuracy.df
plot(accuracy ~ k, data=accuracy.df, type="l", 
     xlab="Number of Neighbors (k)",
     ylab="Accuracy (%)")
```

Find most accurate K for just exit velocity and launch angle on no minimum salary dataset

```{r}
# initialize a data frame with two columns: k, and accuracy.
accuracy.df <- data.frame(k = seq(1, 14, 1), accuracy = rep(0, 14))

# compute knn for different k on validation.
for(i in 1:15) {
  nomin.knn.pred <- knn(train.df[, c(25,26)], valid.df[, c(25,26)], 
                  cl = train.df[, 29], k = i)
  accuracy.df[i, 2] <- confusionMatrix(nomin.knn.pred, as.factor(valid.df[, 29]))$overall[1] 
}

accuracy.df
plot(accuracy ~ k, data=accuracy.df, type="l", 
     xlab="Number of Neighbors (k)",
     ylab="Accuracy (%)")
```

Find most accurate K for all variables excluding exit velocity and launch angle on no minimum salary dataset

```{r}
# initialize a data frame with two columns: k, and accuracy.
accuracy.df <- data.frame(k = seq(1, 14, 1), accuracy = rep(0, 14))

# compute knn for different k on validation.
for(i in 1:15) {
  nomin.knn.pred <- knn(train.df[, c(1:24,27:28)], valid.df[, c(1:24,27:28)], 
                  cl = train.df[, 29], k = i)
  accuracy.df[i, 2] <- confusionMatrix(nomin.knn.pred, as.factor(valid.df[, 29]))$overall[1] 
}

accuracy.df
plot(accuracy ~ k, data=accuracy.df, type="l", 
     xlab="Number of Neighbors (k)",
     ylab="Accuracy (%)")
```