---
title: "5MinuteChallenge"
author: "Shane Regel"
date: "3/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Statcast data
https://baseballsavant.mlb.com/leaderboard/custom?year=2019,2018,2017,2016,2015&type=batter&filter=&sort=4&sortDir=desc&min=100&selections=player_age,b_ab,b_total_pa,b_total_hits,b_single,b_double,b_triple,b_home_run,b_strikeout,b_walk,b_k_percent,b_bb_percent,batting_avg,slg_percent,on_base_percent,on_base_plus_slg,xba,xslg,woba,xwoba,xobp,xiso,exit_velocity_avg,launch_angle_avg,sweet_spot_percent,barrel_batted_rate,&chart=false&x=xba&y=xba&r=no&chartType=beeswarme

```{r}
library(tidyverse)
bat <- read_csv("stats.csv")
summary(bat)
```

Import MLB salary data from 2015-2019
https://www.usatoday.com/sports/mlb/salaries/2015/player/all/

```{r}
salaries <- read_csv("MLBsalaries.csv")
summary(salaries)
```

Split name column into first and last

```{r}
names <- data.frame(do.call('rbind', strsplit(as.character(salaries$NAME),' ',fixed=TRUE)))
names <- names[-c(3, 4)]
names(names) <- c("First", "Last")
salaries <- merge(salaries, names, by.x = 0, by.y = 0)
```

Drop unneccessary columns and pitchers from salary data

```{r}
salaries <- salaries[-c(1, 2, 3, 7, 8, 9)]
salaries <- salaries[!(salaries$POS=="SP" | salaries$POS=="RP"),]
```

Merge salary data with statcast data

```{r}
bat <- merge(bat, salaries, by.x=c("first_name", "last_name", "year"), by.y=c("First", "Last", "YEAR"))
bat <- bat[-c(32)]
```

Add team win totals to data set
https://www.baseball-reference.com/leagues/MLB/2019-standings.shtml

```{r}
wins <- read_csv("wins.csv")
bat <- merge(bat, wins, by.x=c("TEAM", "year"), by.y=c("Team", "Year"))
```

Plot homeruns over time

```{r}
ggplot(bat, aes(x = year, y = b_home_run, group = year)) +            
  geom_boxplot()
```

Plot distributions for players for age, exit velocity, launch angle, sweet spot percent, and barrel rate

```{r}
ggplot(bat, aes(x=player_age)) + geom_histogram(color="darkblue", fill="lightblue")
ggplot(bat, aes(x=exit_velocity_avg)) + geom_histogram(color="darkblue", fill="lightblue")
ggplot(bat, aes(x=launch_angle_avg)) + geom_histogram(color="darkblue", fill="lightblue")
ggplot(bat, aes(x=sweet_spot_percent)) + geom_histogram(color="darkblue", fill="lightblue")
ggplot(bat, aes(x=barrel_batted_rate)) + geom_histogram(color="darkblue", fill="lightblue")
```

Plot exit velocity, launch angle, sweet spot percent, and barrel rate against home runs

```{r}
ggplot(bat, aes(x = exit_velocity_avg, y = b_home_run)) +
    geom_point(aes(color = factor(player_age)))
ggplot(bat, aes(x = launch_angle_avg, y = b_home_run)) +
    geom_point(aes(color = factor(player_age)))
ggplot(bat, aes(x = sweet_spot_percent, y = b_home_run)) +
    geom_point(aes(color = factor(player_age)))
ggplot(bat, aes(x = barrel_batted_rate, y = b_home_run)) +
    geom_point(aes(color = factor(player_age)))
```


Plot exit velocity, launch angle, sweet spot percent, and barrel rate over time since statcasts release

```{r}
ggplot(bat, aes(x = year, y = exit_velocity_avg, group = year)) +            
  geom_boxplot()
ggplot(bat, aes(x = year, y = launch_angle_avg, group = year)) +            
  geom_boxplot()
ggplot(bat, aes(x = year, y = sweet_spot_percent, group = year)) +            
  geom_boxplot()
ggplot(bat, aes(x = year, y = barrel_batted_rate, group = year)) +            
  geom_boxplot()
```


```{r}
ggplot(salaries, aes(x=SALARY)) + geom_histogram(color="darkblue", fill="lightblue")
```

Salary vs Age

```{r}
ggplot(bat, aes(x = player_age, y = SALARY)) +
    geom_point()
```


Plot home runs vs salaries and home runs vs wins

```{r}
ggplot(bat, aes(x = SALARY, y = b_home_run)) +
    geom_point(aes(color = factor(player_age)))
ggplot(bat, aes(x = Wins, y = b_home_run)) +
    geom_point(aes(color = factor(player_age)))
```

Plot launch angle and exit velocity vs salaries and wins

```{r}
ggplot(bat, aes(x = SALARY, y = exit_velocity_avg)) +
    geom_point(aes(color = factor(player_age)))
ggplot(bat, aes(x = Wins, y = exit_velocity_avg)) +
    geom_point(aes(color = factor(player_age)))
ggplot(bat, aes(x = SALARY, y = launch_angle_avg)) +
    geom_point(aes(color = factor(player_age)))
ggplot(bat, aes(x = Wins, y = launch_angle_avg)) +
    geom_point(aes(color = factor(player_age)))
```

