---
title: "Classroom Assignment 5"
author: "Shane Regel"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
In this exercise you will build ensembles, using some in-built commands.


Load file UniversalBank.csv
```{r}
unibank <- read.csv("UniversalBank.csv")
```

Drop ID and zip code columns (1 and 5).
```{r}
unibank <- unibank[,-c(1,5)]
```

Transform Personal.Loan into categorical variable (factor variable)
```{r}
unibank$Personal.Loan <- as.factor(unibank$Personal.Loan)
```

Partition the data 60/40. Set seed at 1
```{r}
set.seed(1)
train_rows <- sample(row.names(unibank), length(unibank$Age)*0.6)
test_rows <- setdiff(row.names(unibank), train_rows)
train_data <- unibank[train_rows,]
test_data <- unibank[test_rows,]
```

Q.1. Train a decision tree using train() function of "caret" package. Use method = "rpart".
Predict variable "Personal.Loan" using rest of the variables.
```{r}
library(caret)
dtree_model <- train(Personal.Loan~., data = train_data, method = 'rpart')
```


Q.2. Predict using predict() function, use type = "raw" to predict the class (not the class probability).
```{r}
pred <- predict(dtree_model, test_data)
```


Q.3. Create confusion matrix using confusionMatrix(prediction, true_value). Use positive = "1" to assign "1" as the positive class.
```{r}
confusionMatrix(pred, as.factor(test_data$Personal.Loan))
```


Q.4. Run bagging algorithm using method = "treebag" in the train() function. Predict Personal.Loan using rest of the variables
```{r}
bag_model <- train(Personal.Loan~., data = train_data, method = 'treebag')
```


Q.5. Predict test data using predict() function. Use type = "raw"
```{r}
pred_bag <- predict(bag_model, test_data, type="raw")
```


Q.6. Obtain confusion matrix using confusionMatrix(prediction$class, true_value). Use positive = "1" to assign "1" as the positive class.
```{r}
confusionMatrix(as.factor(pred_bag), as.factor(test_data$Personal.Loan), positive="1")
```


Q.7. Run boosting algorithm using method = "adaboost" in train() function. 
```{r}
#install.packages("fastAdaboost")
library(fastAdaboost)
boost_model <- train(Personal.Loan~., data=train_data, method="adaboost")
```


Q.8. Predict test data using predict().
```{r}
pred_boost <- predict(boost_model, test_data)
```

Q.9. Create confusion Matrix.
```{r}
confusionMatrix(as.factor(pred_boost), as.factor(test_data$Personal.Loan), positive="1")
```


